Deriving Multi-Headed Planar Dependency Parses from Link Grammar Parses

### 2014-12-13 doc/slides/hong+eisner.TLT13.slides.tex
Changed the slide discussing the reversed subject links
Edited the slides with about half of Jasons suggestions. The other half still coming.

\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{longtable}

@@

  \begin{figure}
    \begin{dependency}
      \begin{deptext}
        Jill \& persuaded \& Jack \& to \& skip \\
      \end{deptext}
      \deproot[edge above, thick, hide label, edge unit distance = 1.5ex]{2}{}
      \depedge[edge above, thick, hide label]{2}{1}{}
      \depedge[edge above, thick, hide label]{2}{3}{}
      \depedge[edge above, thick, hide label]{3}{4}{}
      \depedge[edge above, thick, hide label]{4}{5}{}
      \depedge[edge below, ultra thick, hide label, edge style = {purple}, edge unit distance = 1.5ex]{3}{5}{}
    \end{dependency}
    \caption*{Jack is the object of one verb and the subject of another}
  \end{figure}
}

@@

\frame{\frametitle{Link Grammars}
  \begin{itemize}
    \item[] Grammar-based formalism for projective dependency parsing with undirected links
  \end{itemize}
}

@@

Encoded Constraints:
\begin{itemize}
  \item Connectedness
  \item Acyclicity
  \item Consistency of Directionalized Links
\end{itemize}
}

@@

Variables: 
\begin{itemize}
\item[] $x_{ij}, x_{ji} \in \mathbb{Z} \geq 0$: orientation of each link
\item[] $x_{ij} + x_{ji} = 1$

%\item $x_{ij} + x_{ji} = 1$ (A link can only be oriented left or right)

\end{itemize}

%\NoteJH{TODO: Incomplete slide. Need to describe things like depth of the variable, label}
}

\frame{\frametitle{Connectedness, Acyclicity}
\begin{itemize}
  \item[] Connectedness
    \begin{align}
      \sum_u x_{uv} & \geq 1
    \end{align}

@@

    \begin{align}
      (\forall_u)\; d_v + (1 + n_v) \cdot (1 - x_{uv}) & \geq 1+d_u
    \end{align}
\end{itemize}

@@

\frame{\frametitle{Stability of Results}
  \begin{figure}
    \centering
    \includegraphics[width=0.7\linewidth]{precision_recall}
  \end{figure}
}

@@

  \uncover<2->{Multiheadedness}
  \begin{itemize}
  \uncover<2->{\item[] Link Data has 8\% additional edges over the CoNLL.}
  \end{itemize}

@@

\begin{itemize}
\item In our paper we claimed this was due to an inconsistency of the Link Grammar, discovered by our method.

\end{itemize}

\begin{figure}
  \begin{dependency}
	\begin{deptext}
	  Jill \& wanted \& to \& skip \\
	\end{deptext}
    \deproot[edge above, edge unit distance=2ex]{1}{W}
    \deproot[edge above, edge unit distance=2ex]{2}{WV}
	\depedge[edge above]{2}{1}{S}
	\depedge[edge above]{2}{3}{TO}
	\depedge[edge above]{3}{4}{I}
	\depedge[edge above]{2}{4}{IV}
  \end{dependency}
\end{figure}

%\begin{figure}
%  \begin{dependency}
%	\begin{deptext}
%	  Jill \& persuaded \& him \& to \& skip \\
%	\end{deptext}
%    \deproot[edge above, edge unit distance=2ex]{1}{W}
%    \deproot[edge above, edge unit distance=2ex]{2}{WV}
%	\depedge[edge above]{2}{1}{S}
%	\depedge[edge above]{2}{3}{O}
%	\depedge[edge above]{2}{4}{TO}
%	\depedge[edge above]{4}{5}{I}
%	\depedge[edge above]{2}{5}{IV}
%  \end{dependency}
%\end{figure}

\begin{figure}
  \begin{dependency}
	\begin{deptext}
	  Jill \& thinks \& he \& will \& skip \\
	\end{deptext}
    \deproot[edge above, edge unit distance=2ex]{1}{W}
    \deproot[edge above, edge unit distance=2ex]{2}{WV}
	\depedge[edge above]{1}{2}{S}
	\depedge[edge above]{2}{3}{C}
	\depedge[edge above]{3}{4}{S}
	\depedge[edge above]{4}{5}{I}
	\depedge[edge above]{2}{5}{CV}
  \end{dependency}
\end{figure}

@@

>>>>

\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{longtable}
\usepackage{pdfpages}

@@

  \begin{figure}
    \begin{dependency}
      \begin{deptext}
        Jill \& persuaded \& Jack \& to \& skip \\
      \end{deptext}
      \deproot[edge above, thick, hide label, edge unit distance = 1.5ex]{2}{}
      \depedge[edge above, thick, hide label]{2}{1}{}
      \depedge[edge above, thick, hide label]{2}{3}{}
      \depedge[edge above, thick, hide label]{3}{4}{}
      \depedge[edge above, thick, hide label]{4}{5}{}
      \depedge[edge below, ultra thick, hide label, edge style = {purple}, edge unit distance = 1.5ex]{5}{3}{}
      \depedge[edge below, ultra thick, hide label, edge style = {purple}, edge unit distance = 1.5ex]{2}{5}{}
    \end{dependency}
    \caption*{Jack is the object of one verb and the subject of another}
  \end{figure}
}

@@

\frame{\frametitle{Link Grammars}
  \begin{itemize}
  \item[] Grammar-based formalism for projective dependency parsing with undirected links
  \item[] Orignal formalism and English Link Grammar created by Davy Temperley, Daniel Sleator, John Lafferty (1991)
  \end{itemize}
}

\subsection{Link Grammars Intro}
\frame{\frametitle{Link Grammars}
  \includegraphics[width=\linewidth]{linkgrammar_example1.png}\footnote{These figures were clipped from the orginal Link Grammar paper: \\``Parsing English with a Link Grammar'' by Sleator and Temperley}
}

\frame{\frametitle{Link Grammars}
  \includegraphics[width=\linewidth]{linkgrammar_example2.png}
}

\frame{\frametitle{Link Grammars}
  \includegraphics[width=\linewidth]{linkgrammar_example3.png}
}

@@

\subsection{What is ILP?}
\frame{\frametitle{What is Integer Linear Programming?}
\begin{itemize}
  \item Its an optimization problem where some or all of the variables are integers. 
  \item The objective function and constraints are linear.
  \item Its NP-Hard! But good solvers exist that work well most of the time.
  \item Our ILP is encoded as a ZIMPL program and solved using the SCIP Optimization Suite
\end{itemize}
}

\subsection{Corpus Building}
\frame{\frametitle{Corpus Building Strategy}
\begin{itemize}
  \item We start with a corpus of link grammar parses.
  \item We try to directionalize this corpus, choosing a direction for every link in every parse.
  \item We use an Integer Linear Program to find consistent directions for every link
\end{itemize}
}

\subsection{ILP Model}
\frame{\frametitle{Integer Linear Programming Model}
Encoded Constraints:
\begin{itemize}
  \item Connectedness\uncover<2->{: Every word has a link attaching it}
  \item Acyclicity\uncover<2->{: No cycles!}
  \item Consistency of Directionalized Links\uncover<2->{:

    Similar links oriented the same way}
\end{itemize}
}

@@

Variables: 
\begin{itemize}
\item[] $x_{ij}, x_{ji} \in \mathbb{Z} \geq 0$: orientation of each link
\item[] $x_{ij} + x_{ji} = 1$
\end{itemize}
\uncover<2->{\begin{center}
    A link can only be oriented left or right
\end{center}}

\frame{\frametitle{Connectedness, Acyclicity}
\begin{itemize}
  \item[] Connectedness
    \begin{align}
      \sum_u x_{uv} & \geq 1 
    \end{align}
    \uncover<2->{\begin{center}
        A word has at least 1 parent
    \end{center}}

@@

    \begin{align}
      (\forall_u)\; d_v + (1 + n_v) \cdot (1 - x_{uv}) & \geq 1+d_u
    \end{align}
    \uncover<2->{\begin{center}
        The depth of a child is greater than the depth of the parent
    \end{center}}
\end{itemize}

@@

\frame{\frametitle{Stability of Results}

  \begin{itemize}
  \item We worried that the recovered direction mapping might be unstable and sensitive to the input corpus. 
  \item We compared the results of increasing runs of sentences.
  \end{itemize}

  \begin{figure}
    \centering
    \includegraphics[width=0.6\linewidth]{precision_recall}
  \end{figure}
}

@@

  \uncover<2->{Multiheadedness}
  \begin{itemize}
  \uncover<2->{\item[] Link Data has 8\% additional edges over the CoNLL.

    (average about 2 multiheaded words per sentence)}
  \end{itemize}

@@

\begin{itemize}
\item This is due to a possible inconsistency of the Link Grammar, discovered by our method. 
\end{itemize}

%    +-------WV-------+--------CV-------+
%    +---Wd---+---Ss--+--Ce-+-Ss-+---I--+
%    |        |       |     |    |      |
%LEFT-WALL Jill.f thinks.v he will.v skip.v

\begin{figure}
\vspace{-6pt}
\centering
  \begin{dependency}[edge style={-}]
	\begin{deptext}
	  Jill \& thinks \& he \& will \& skip\\
	\end{deptext}
    %\deproot[edge above, edge unit distance=2ex]{1}{W}
    %\deproot[edge above, edge unit distance=2ex]{2}{WV}
	\depedge[edge above]{1}{2}{S}
	\depedge[edge above]{2}{3}{C}
	\depedge[edge above]{3}{4}{S}
	\depedge[edge above]{4}{5}{I}
	\depedge[edge above]{2}{5}{CV}
  \end{dependency}
  %\caption*{Links with label I is forced to go right}
\end{figure}

%    +--------------WV-------------+
%    +---Wd---+---Ss--+--MVi-+--I--+
%    |        |       |      |     |
%LEFT-WALL Jill.f thinks.v to.r skip.v

\begin{figure}
\vspace{-6pt}
\centering
  \begin{dependency}[edge style={-}]
	\begin{deptext}
	  Jill \& hopes \& to \& skip\\ % Originally: Jill thinks to skip. The sentence was altered to be clearer.
	\end{deptext}
    %\deproot[edge above, edge unit distance=2ex]{1}{W}
    %\deproot[edge above, edge unit distance=2ex]{2}{WV}
	\depedge[edge above]{1}{2}{S}
	\depedge[edge above]{2}{3}{MV}
	\depedge[edge above]{3}{4}{I}
  \end{dependency}
  %\caption*{Links with label I is forced to go right}
\end{figure}

@@

%\frame{\frametitle{Possible simpler explanation}

%linkparser> 
%linkparser> Jill made him go
%Found 5 linkages (3 had no P.P. violations)
%Linkage 1, cost vector = (UNUSED=0 DIS= 0.00 LEN=6)
%
%    +-------WV-------+----I*j---+
%    +---Wd---+---Ss--+--Ox-+    |
%    |        |       |     |    |
%LEFT-WALL Jill.f made.v-d him go.v

%\begin{figure}
%\vspace{-6pt}
%\centering
%  \begin{dependency}
%	\begin{deptext}
%	  Jill \& made \& him \& go \\
%	\end{deptext}
%    \deproot[edge above, edge unit distance=2ex]{1}{W}
%    \deproot[edge above, edge unit distance=2ex]{2}{WV}
%	\depedge[edge above]{1}{2}{S}
%	\depedge[edge above]{2}{4}{I}
%	\depedge[edge above]{2}{3}{O}
%  \end{dependency}
%  \caption*{Links with label I is forced to go right}
%\end{figure}

%\begin{figure}
%  \vspace{-12pt}
%  \begin{dependency}
%	\begin{deptext}
%	  Jill \& would \& go \\
%	\end{deptext}
%    \deproot[edge above, edge unit distance=2ex]{1}{W}
%    \deproot[edge above, edge unit distance=2ex]{3}{WV}
%	\depedge[edge above]{1}{2}{S}
%	\depedge[edge above]{2}{3}{I}
%  \end{dependency}
%  \caption*{If I is forced to go right, then S is as well}
%\end{figure}
%}
